---
title: "R file size issues"
author: "Keaton Stagaman"
date: "1/8/2021"
output: github_document
---

Below reads in a phyloseq object with a very large community matrix (47 x 1,060,929)

```{r setup, echo=TRUE}
library(phyloseq)

obj.size <- function(x) {paste("Object size:", format(object.size(x), units = "auto"))}
f.size <- function(x) {paste("File size:", round(file.size(x) / 1024^2, 1), "Mb")}

ps <- readRDS("example_phyloseq.rds")
```

Next, I create a distance matrix to determine community dissimilarities between samples
It should result in a 47 x 47 matrix

```{r create-distance-matrix, echo=TRUE}
dist.mat <- phyloseq::distance(ps, method = "bray") # creating a distance matrix
attributes(dist.mat)$Size
```

Okay, as expected, but...
```{r issue1}
cat(obj.size(dist.mat), sep = "\n")
```

This is oddly big for a distance matrix of this size (like orders of magnitude larger)

After investigating I realize that this is because `attributes(dist.mat)$call` is huge, because it includes every single value from the original community matrix.

```{r possible-solution}
dist.mat.small <- dist.mat
attributes(dist.mat.small)$call <- NULL
cat(obj.size(dist.mat.small), sep = "\n")
```

Okay, so that's looking good. Much closer to what I'd expect for the size. But...when I try to save...

```{r issue3}
saveRDS(dist.mat, file = "distance_matrix_original.rds")
saveRDS(dist.mat.small, file = "distance_matrix_small.rds")

cat(f.size("distance_matrix_original.rds"), sep = "\n")
cat(f.size("distance_matrix_small.rds"), sep = "\n")
```

What's happening here? Why are the file sizes the same for objects of very different size?

```{r cleanup, include=FALSE}
file.remove("distance_matrix_original.rds")
file.remove("distance_matrix_small.rds")
```

